

# Exercises and Quizzes

## Exercises {#Sec:Exercises}

### Incorporating Exercises into the Textbook

Users expect to see exercises in a textbook. For *Loss Data Analytics*, we have many sources and types of exercises that we can draw upon. The issue is deciding upon the type to re-inforce the emphasis of the book.

Currently, the book features almost as many different exercise types as there are authors. Some chapters have no exercises, others have an extensive amount. (See also some questions in a [Future Work](https://github.com/ewfrees/InteractiveLDA/tree/master/FutureWork) folder on this site.)

These differing viewpoints are also reflected in the examples in the book. To illustrative, some chapters focus on mathematical/statistical developments, some on data analytic  aspects, some on business content. This is the good and bad associated with a heterogenous collection of authors.

We can reduce the effects of this heterogeneity by providing a large pool of exercises. Some may exist outside of the book. With a large pool of exercises, users (teachers) can draw upon the subset that reflects their area of interests.

### Wordpress Exercises

Here is a series of exercises that guide the viewer through some of the theoretical foundations of *Loss Data Analytics*. Each tutorial is based on one or more questions from the professional actuarial examinations – typically the Society of Actuaries short term actuarial mathematics exam that used to be known as *Exam C*. These exercises are coded using **Wordpress**. 

Based on Google Analytics data, this approach is popular. If we want to continue to develop this line of throught, one approach is to explore bringing these exercises into the `R` **Bookdown** environment.

#### Frequency Modeling

<p style="text-align: center;">
[Frequency Distribution Guided Tutorials](https://www.ssc.wisc.edu/~jfrees/loss-data-analytics/loss-data-analytics-problems/)
</p>

#### Modeling Loss Severity

<p style="text-align: center;">
[Severity Distribution Guided Tutorials](http://www.ssc.wisc.edu/~jfrees/loss-data-analytics/chapter-3-modeling-loss-severity/loss-data-analytics-severity-problems/)
</p>

### Incorporating Datacamp Exercises into the Short Course

Describe short course exercises here ...

There are a variety of ways of learning coding procedures interactively. One option is [Datacamp](https://www.datacamp.com). In particular, they have developed an `R` package, [Datacamp Light](https://support.datacamp.com/hc/en-us/articles/360007749853-What-is-DataCamp-Light-). This allows us to integrate this interactive approach into our authoring system. [Here](https://ewfreesres.github.io/RegressModel/index.html) is a link to an online regression tutorial done using this approach. There are a few little technical details but we think these are now ironed out. This is a promising option that will help make the book special.

## End of Section Quizzes {#Sec:EndSectionQuizzes}

The more straight-forward aspect is the inclusion of end of the section quizzes. Unlike exercises, these are designed to be low-level, formative assessment tools that a person can use to assess whether there is some comprehension of material just read. Here are two end of section quizzes from Chapter 1, just to provide a flavor. More are available in [*Loss Data Analytics*](https://openacttexts.github.io/Loss-Data-Analytics/index.html).

### Section 1.1 

```{r child = './Quizzes/Quiz11.html'} 
```

### Section 1.2

```{r child = './Quizzes/Quiz12.html'}
```


